Weight Lifting Form Prediction using Accelerometer Data
========================================================

# Introduction

The goal of his project is to predict the weight lifting form from the accelerometer data collected from 6 participants, who were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
We use the random forest model which provides very good prediction of the form of the weight lifting exercises.


# Loading and Processing the data
Loading libraries and reading in the training and testing sets from 
our csv files. 

```{r}
library(RANN)
library(caret)
training_raw <- read.csv('pml-training.csv')
testing_raw <- read.csv('pml-testing.csv')
```

Splitting data into a training and testing set. We will split the data 
into 70% for the training set and 30% for the testing set. 
```{r}
subset <- createDataPartition(training_raw$classe, list=FALSE, p=.7)

training = training_raw[subset,]
testing = training_raw[-subset,]
```


Filter the numeric features and outcome, using the PreProcess
function to deal with the missing variables/data from the set.
We then rename the first variable to classe.
```{r}
n = which(lapply(training, class) %in% c('numeric')  )

preModel <- preProcess(training[,n], method=c('knnImpute'))

training2 <- cbind(training$classe, predict(preModel, training[,n]))
names(training2)[1] <- 'classe'

testing2 <- cbind(testing$classe, predict(preModel, testing[,n]))
names(testing2)[1] <- 'classe'
```

Now we create a test set. 
```{r}
test <- predict(preModel, testing_raw[,n])
```

The 5 levels of the variable Classe are:     
**Class A**: exactly according to the specification    
**Class B**: throwing the elbows to the front    
**Class C**: lifting the dumbbell only halfway   
**Class D**: lowering the dumbbell only halfway   
**Class E**: throwing the hips to the front  

# Random Forest Model
We will now use the randomForest dataset to fit a random tree model.
We also set up our confusion matrix.

```{r, warning=FALSE}
library(randomForest)
forest  <- randomForest(classe ~ ., training2)

training_pred <- predict(forest, training2) 
print(confusionMatrix(training_pred, training2$classe))
```
The in-sample accuracy level is 100%.

# Out-of-sample 
```{r}
out_of_sample <- predict(forest, testing2) 
```

Confusion Matrix: 
```{r}
print(confusionMatrix(out_of_sample, testing2$classe))
```

The cross validation accuracy is  98.6%. 
The model is performing quite well and we can keep this model to predict the classe variable on the test set.

# Test Set Prediction Results

Applying model to the test data.
```{r}
results <- predict(forest, test) 
results
```

The model achieved a high accuracy rate in and out of samples and correctly predicted the 20 instances of the test set.
We now print our output into separate txt files.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_",i,".txt")
    write.table(x[i], file = filename, quote=FALSE, row.names=FALSE, col.names=FALSE)
  }
}

pml_write_files(results)
```
